\lecture{12}{2025 February 24 15:58}{4th Declension}
If you recall last time we mentioned that there are 5 groups
of nouns called Declension \\ \\ 
Fourth declension displays a characteristic -u- throught 
their declension (except in the dative and ablative plural).
They are also identified by the -us- in the genitive singular.
\begin{center}  
  \captionof{table}{4th Declension}
  \begin{tabular}{lll}
    senatus, senatus (M.) Senate  \\
    & Singular & Plural \\ 
    Nom. & senatus & senatus \\ 
    Genitive & senatus & senatuum \\ 
    Dative & senatui & senatibus \\ 
    Accusative & senatum & senatus \\ 
    Ablative & senatu & senatibus 
  \end{tabular}

\end{center}
\begin{center}
  \huge Cicero VS. Catiline \\ Latin
\end{center}
O tempora O mores! Senatus haec intellegit, consul videt; hic
tamen vivit. Vivit? Immovero etiam in senatum venit, fit publici
consili particepts, notat et designat oculis adcaedem unum quemque
nostrum. Nos autem fortes viri satis facere rei publicae videmur,
si istius furorem ac tela vitamus.
\begin{center}
  \huge Cicero VS. Catiline \\ English
\end{center}
O such times, such morals! The senate is aware of these things;
the consul sees them; and yet this man lives. He lives? He even 
comes into the senate and participates in the public deliberations; 
he is watching and marking down and checking off for execution 
every individual among us. And we, gallant men that we are, think that 
we are doing our duty to the Republic if we avoid his fury and attacks
