\lecture{9}{2025 February 17 16:46}{1st Declension}
At first learning latin can often seem intimidating especially vocabulary
that's why we tend to clasify nouns into five major groups and we call these
groups Declensions \\\\ 
The first declension used to be called the femine because most of the nouns
belonging to this class were associated with femine gendered nouns like puellla
or femina \\\\ 
The fact is some masculine nouns such as agricola and nauta also share these
endings and they are certainly masculine

\begin{center}  
  \captionof{table}{1st Declension}
  \begin{tabular}{lllll}
    \\
    puella -ae (F.) girl & & & \\
    & Singular & & Plural &  \\
    Nom. & puella & the girl & puellae & the girls \\
    Genitive & puellae & of the girl & puellarum & of the girls \\ 
    Dative & puellae & to the girl & puellis & to the girls \\
    Accusative & puellam & the girl & puellas & the girls \\ 
    Ablative & puella & by the girl & puellis & by the girls \\ 
  \end{tabular}
\end{center}
Lets put what we know into practice:\\\\
\begin{tabular}{lll}
  & English & Latin & 
  Subject (Nom.) & The farmer is famous & Agricola est fama \\ 
  Possesion (Genitive) & The life of the farmet & Vita agricolae \\ 
  Indirect Object (Dative) & The farmer gives a note to her & Agricola notam eae dat \\ 
  Direct Object (Accusative) & The sailor drinks water & Nauta aquam bibit \\ 
  Object of the preposition (Abl.) & The sailor sails by the stars & Nauta stellis navigat \\ 
  Direct Address (Vocative) & O Goddess & O Deae \\
\end{tabular}
Now lets us try some sight translation \\ \\ 
\begin{center}
 \huge Pandoras's Box \\
 Latin
\end{center}
Pandorra est femina. Pandora semper est curiosa. Uno die arcam accipit. 
Nota dicit: non aperi! Suus vir, Epimetheus, eam non aperi monet. Sed,  
eam aperit. Miseriae effungiunt. Spes tamen remanet. 
 \begin{center}
 \huge Pandoras's Box \\
 English
\end{center}
Pandora is a woman. Pandora is always curious. One day, she receives a box. 
The note says: do not open. Her husband,Epimetheus, warns her to not open it. 
But she opens it. All the troubles escapce. Nevertheless, Hope remains.
